Data preparation

### 2024-05-08 paper1_jiturra.Rmd
update abstract y class inequality

---
title: "Class-based network segregation, Economic Inequality and Redistributive Preferences across societies"
css: "input/css/custom.css" # custom CSS para html
fontsize: 12pt
linestretch: '1.15'          # interlineado 
papersize: a4
link-citations: yes         # citas linkeadas
lang: en-US
# date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author:
- name: Julio Iturra-Sanhueza
  affiliation: Bremen International Graduate School of Social Sciences
  email: jiturra@uni-bremen.de
#   number: 1
# - name: Justin d'Ottawa
#   affiliation: University of Ottawa
#   number: 2
# - name: Pedro Torres
#   number: 1,2 
# Nota: Autores que comparten filiacion, poner el mismo number y declarar filiación una sola vez.  
abstract: | 
  | The escalating economic inequality and the current sanitary crisis have threatened social solidarity among citizens. However, these consequences have been unevenly experienced across the social structure where the upper and intermediate classes, benefiting from privileged access to social resources, have shown resilience. Conversely, working-class families have faced deteriorating material conditions, leading to a heightened sense of marginalization and increasing the demand for welfare support. Particularly, social resources through network ties have been suggested to have a direct link with an individual's welfare through instrumental and expressive outcomes, such as providing information or help in moments of need. Consequently, it has been argued that being socialy excluded not only plays a role in terms of social resources, but higher levels of segregation can bolster opinions, suggesting that being isolated from other social classes can polarise attitudes mainly in the working and upper-middle classes. In this regard, based on the ISSP 2017 -  Social Networks (N=31,191),  multilevel estimations suggest that the influence of network homogeneity is conditional to social class, where the working and intermediate classes with high homogeneous networks hold higher redistributive preferences than the upper classes. Additionally, income inequality has been shown to decrease the polarizing effect of network segregation, mainly among the upper classes, when overall inequality is high. In sum, this study highlights the role of both social class and network segregation together to understand redistributive preferences, as well as their implications for policies and cohesion.
  | 
  | **Keywords**: network segregation, redistributive preferences, economic inequality
output:
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    number_sections: true
    pandoc_args:
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion  
  bookdown::html_document2:
    toc: yes
    code_folding: hide
    number_sections: false
    toc_float:
      collapsed: false
      smooth_scroll: false

@@

It has been previously established that wider economic gaps lead to segregated forms of social participation and status composition of social networks [@pichler_patterns_2007;@otero_space_2022]. Indeed, people with lower economic backgrounds have fewer chances to participate in social activities and form diverse social networks. Therefore, it has been suggested that material inequality reinforces the stratified access to social activities and the social distance between social classes, resulting in an increasing marginalization of people with lower resources and consolidating the privileged positions of the upper classes that held better opportunities for diversifying their social activities and interpersonal ties [@lancee_income_2012;@letki_getting_2015;@otero_differences_2023].

@@

As social relations within the occupational structure serve as the foundation for class analysis, widening our examination to other dimensions of social life, such as family dynamics, the development of friendships, and diverse social activities, can significantly enrich the scope of class analysis, particularly concerning its relevance in political attitudes formation. In principle, social stratification scholars have traditionally considered differentiation as observable patterns of association given by the intersection of status and group membership, which can be understood as relational networks between positions [@blau_inequality_1977]. Empirically, homophily is one of the most significant findings in network research, understood as the higher likelihood of tie formation between people of similar characteristics [@mcpherson_birds_2001, p.416] that trace friendship and family ties, such as age and gender, as well as segregated social environments alongside ethnicity and socioeconomic status [@bargsted_pautas_2020; @plaza_juntos_2022;@lazarsfeld_friendship_1954]. In addition, psychosocial factors such as socialization preferences and group dynamics also play a role in segregation processes, considering that isolated social environments decrease the interchange of worldviews, reinforcing in-group selectivity and value similarity [@visser_attitudes_2004;@homans_human_1951]. Nonetheless, without denying the relevance of socialization preferences, segregation has been predominantly explained through social differentiation in social activities as foci of contact opportunities that ultimately consolidate social networks [@feld_focused_1981; @mcpherson_network_2019]. From this perspective, we argue that class relations represent not only resource-based distinctions but also patterns of sociability and exchange observed in the differentiation of social ties.

@@

Our theoretical approach has established that social classes should be understood beyond labor market relationships but as relational networks that can provide an explanatory framework for attitude formation.
Theoretically, what the literature has established as processes of social influence [@lindh_missing_2021] or political socialization [@lee_consider_2023] has echoed in how social closure in interpersonal networks generates distance between different social classes in terms of their redistributive preferences. Here, the extent to which social classes distance themselves in terms of network segregation indeed creates an "empathy gulf" [@sachweh_moral_2012] that can be observed in how the service class is less willing to take action to strive with inequality as a matter of collective commitment when they are highly segregated in homogeneous upper-class environments [@otero_power_2023], in contrast to the increasing redistributive pressures of the marginalized working class. Besides, as network homogeneity represents low cross-class ties, it can be argued that lower social integration [@blau_inequality_1977] can undermine social solidarity as it consolidates weaker chances of contact between social classes and the lack of knowledge about the lives of others [@vargassalfate_contact_2023].

>>>>

---
title: "Class-based network segregation, Economic Inequality and Redistributive Preferences across societies"
css: "input/css/custom.css" # custom CSS para html
fontsize: 12pt
linestretch: '1.15'          # interlineado 
papersize: a4
link-citations: yes         # citas linkeadas
lang: en-US
# date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author:
- name: Julio Iturra-Sanhueza
  affiliation: Bremen International Graduate School of Social Sciences
  email: jiturra@uni-bremen.de
#   number: 1
# - name: Justin d'Ottawa
#   affiliation: University of Ottawa
#   number: 2
# - name: Pedro Torres
#   number: 1,2 
# Nota: Autores que comparten filiacion, poner el mismo number y declarar filiación una sola vez.  
abstract: | 
  | Escalating economic inequality has increased attention to the link between class-based social ties and redistributive preferences in contemporary societies. However, social classes are not equally affected by inequality regarding their chances of establishing cross-class social relations. While those in the upper and middle classes have greater access to economic and diverse social ties, the working class has been disproportionately affected by social exclusion. Consequently, this leads to a growing divide between their lives and the lives of others, consolidating segregated lifeworlds. Previous research has shown that the link between class relations and redistributive preferences can be enriched from a network perspective. Meeting people from diverse class positions can help individuals become more aware of the lifestyles and worldviews of others. On the one hand, when upper-class networks become more homogeneous, it can lead to lower empathy toward those in need and less support for redistributive policies. On the other hand, when there is segregation in lower-class social networks, it can lead to greater marginalization and a higher demand for welfare support. This study explores how class-based network segregation influences redistributive preferences from a cross-national perspective, employing data from the International Social Survey 2017, which comprised 32,529 individuals in 31 societies. The regression analyses indicate that network homogeneity has a conditional influence on redistributive preferences based on social class. Homogeneous social networks in the working and intermediate classes increase redistributive preferences, while greater homogeneity in the upper classes reduces support for redistribution. In addition, in contexts of greater income inequality, the conditional influence of network segregation weakens, especially for the upper classes. Implications for the study of class relations and policy preferences are discussed.
  | 
  | **Keywords**: social networks, social class, redistributive preferences, income inequality
output:
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    number_sections: true
    pandoc_args:
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion  
  bookdown::html_document2:
    toc: yes
    code_folding: hide
    number_sections: false
    toc_float:
      collapsed: false
      smooth_scroll: false

@@

It has been previously established that wider economic gaps lead to segregated forms of social participation and status composition of social networks [@pichler_patterns_2007;@otero_space_2022]. Indeed, it has been suggested that material inequality reinforces the stratified access to social activities and the social distance between social classes, resulting in an increasing marginalization of people with lower resources and consolidating the privileged positions of the upper classes that held better opportunities for diversifying their social activities and interpersonal ties [@lancee_income_2012;@letki_getting_2015;@otero_differences_2023].

@@

As social relations within the occupational structure serve as the foundation for class analysis, widening our examination to other dimensions of social life, such as family dynamics, the development of friendships, and diverse social activities, can significantly enrich the scope of class analysis, particularly concerning its relevance in political attitudes formation. In principle, social stratification scholars have traditionally considered differentiation as observable patterns of association given by the intersection of status and group membership, which can be understood as relational networks between positions [@blau_macrosociological_1977]. Empirically, homophily is one of the most significant findings in network research, understood as the higher likelihood of tie formation between people of similar characteristics [@mcpherson_birds_2001, p.416] that trace friendship and family ties, such as age and gender, as well as segregated social environments alongside ethnicity and socioeconomic status [@bargsted_pautas_2020; @plaza_juntos_2022;@lazarsfeld_friendship_1954]. In addition, psychosocial factors such as socialization preferences and group dynamics also play a role in segregation processes, considering that isolated social environments decrease the interchange of worldviews, reinforcing in-group selectivity and value similarity [@visser_attitudes_2004;@homans_human_1951]. Nonetheless, without denying the relevance of socialization preferences, segregation has been predominantly explained through social differentiation in social activities as foci of contact opportunities that ultimately consolidate social networks [@feld_focused_1981; @mcpherson_network_2019]. From this perspective, we argue that class relations represent not only resource-based distinctions but also patterns of sociability and exchange observed in the differentiation of social ties.

@@

Our theoretical approach has established that social classes should be understood beyond labor market relationships but as relational networks that can provide an explanatory framework for attitude formation.
Theoretically, what the literature has established as processes of social influence [@lindh_missing_2021] or political socialization [@lee_consider_2023] has echoed in how social closure in interpersonal networks generates distance between different social classes in terms of their redistributive preferences. Here, the extent to which social classes distance themselves in terms of network segregation indeed creates an "empathy gulf" [@sachweh_moral_2012] that can be observed in how the service class is less willing to take action to strive with inequality as a matter of collective commitment when they are highly segregated in homogeneous upper-class environments [@otero_power_2023], in contrast to the increasing redistributive pressures of the marginalized working class. Besides, as network homogeneity represents low cross-class ties, it can be argued that lower social integration [@blau_macrosociological_1977] can undermine social solidarity as it consolidates weaker chances of contact between social classes and the lack of knowledge about the lives of others [@vargassalfate_contact_2023].

### 2024-05-08 procesamiento/02-prod_prep-macro.Rmd
update abstract y class inequality

# centering micro level

```{r eval=FALSE, include=FALSE}
df2 <- 
df2 %>% 
  mutate(to_dummy(female),
         to_dummy(union),
         to_dummy(workst),
         # dummy for categorical variables-------------------------------------
         to_dummy(class3),to_dummy(class3spo),to_dummy(class3res),to_dummy(Q03pcm))
df2$female_gc = group_center(df2$female_2, grp = df2$country2)
df2$union_gc = group_center(df2$union_2, grp = df2$country2)
df2$workst_gc = group_center(df2$workst_2, grp = df2$country2)
df2$edyears_gc = group_center(df2$edyears, grp = df2$country2)
df2$agenum_gc = group_center(df2$agenum, grp = df2$country2)

# EGP: dominance
df2$class3_ser_gc = group_center(df2$class3_1, grp = df2$country2)
df2$class3_mid_gc = group_center(df2$class3_2, grp = df2$country2)
df2$class3_wor_gc = group_center(df2$class3_3, grp = df2$country2)
# EGP: Respondent
df2$class3res_ser_gc = group_center(df2$class3res_1, grp = df2$country2)
df2$class3res_mid_gc = group_center(df2$class3res_2, grp = df2$country2)
df2$class3res_wor_gc = group_center(df2$class3res_3, grp = df2$country2)
# EGP: Respondent
df2$class3spo_ser_gc = group_center(df2$class3spo_1, grp = df2$country2)
df2$class3spo_mid_gc = group_center(df2$class3spo_2, grp = df2$country2)
df2$class3spo_wor_gc = group_center(df2$class3spo_3, grp = df2$country2)
# Household Income Tercile
df2$Q03pcm_1_gc = group_center(df2$Q03pcm_1, grp = df2$country2)
df2$Q03pcm_2_gc = group_center(df2$Q03pcm_2, grp = df2$country2)
df2$Q03pcm_3_gc = group_center(df2$Q03pcm_3, grp = df2$country2)
df2$Q03pcm_NA_gc = group_center(df2$Q03pcm_4, grp = df2$country2)

names(df2)
# drop dummies
df2 <- 
  df2 %>% 
  select(-c(female_1,female_2,union_1,union_2,workst_1,workst_2,
            class3_1,class3_2,class3_3,class3res_1,class3res_2,class3res_3,
            class3spo_1,class3spo_2,class3spo_3,
            Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4))
names(df2)
```

# centering macro level

```{r eval=FALSE, include=FALSE}
df2$gini_disp_c <- center(df2$gini_disp)
df2$gini_mkt_c <- center(df2$gini_mkt)
df2$palmaratio_c <- center(df2$palmaratio)
df2$top10_c <- center(df2$top10)
df2$middle50_c <- center(df2$middle50)
df2$d10d1_c <- center(df2$d10d1)
df2$giniindex_c <- center(df2$giniindex)
df2$ttheilge1_c <- center(df2$ttheilge1)
df2$rgdpna_c <- center(df2$rgdpna)

df2$gini_disp_z <- center(df2$gini_disp)
df2$gini_mkt_z <- center(df2$gini_mkt)
df2$palmaratio_z <- center(df2$palmaratio)
df2$top10_z <- center(df2$top10)
df2$middle50_z <- center(df2$middle50)
df2$d10d1_z <- center(df2$d10d1)
df2$giniindex_z <- center(df2$giniindex)
df2$ttheilge1_z <- center(df2$ttheilge1)
df2$rgdpna_z <- center(log(df2$rgdpna))

df2 %>% 
  select(country3,ends_with("_c")) %>% group_by(country3) %>% summarise_all(.funs = mean) %>% View()
table(df2$country2)
```

>>>>

# Class inequality

```{r}
summary(df1$incomepcap)

df_class3_ineq <- 
df2 %>%   
  group_by(country2,dclass3res_V) %>%
  summarise_at(c("incomepcap","ednum","wid_gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame() %>%
  na.omit() %>% 
  dplyr::select(country2,dclass3res_V,incomepcap,ednum,wid_gini_disp) %>% 
  tidyr::pivot_wider(names_from = "dclass3res_V", values_from = c("incomepcap","ednum")) %>% 
  mutate(ratio_class_inc = `incomepcap_Service Class (I+II)`/`incomepcap_Working Class (V+VI+VII)`,
         ratio_class_educ  = `ednum_Service Class (I+II)`/`ednum_Working Class (V+VI+VII)`) %>% 
  as.data.frame()

df_class3_ineq %>% select(wid_gini_disp,ratio_class_inc,ratio_class_educ) %>% tab_corr()

df_class6_ineq <- 
df2 %>%   
  group_by(country2,dclass6res) %>%
  summarise_at(c("incomepcap","ednum","wid_gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame() %>%
  na.omit() %>% 
  dplyr::select(country2,dclass6res,incomepcap,ednum,wid_gini_disp) %>% 
  tidyr::pivot_wider(names_from = "dclass6res", values_from = c("incomepcap","ednum")) %>% 
  mutate(ratio_class_inc = `incomepcap_Upper Service class`/`incomepcap_Unskilled working class`,
         ratio_class_educ  = `ednum_Upper Service class`/`ednum_Unskilled working class`) %>% 
  as.data.frame() 
df_class_ineq %>% select(wid_gini_disp,ratio_class_inc,ratio_class_educ) %>% tab_corr()
```

```{r}

income_data

# Example data
# Assuming you have a data frame called 'income_data' with columns: country, social_class, and average_income

# Calculate all pairs of income ratios within each country
income_ratios <- income_data %>%
  group_by(country) %>%
  mutate(income_ratio = combn(average_income, 2, FUN = function(x) x[1] / x[2])) %>%
  ungroup()

# Print the result
print(income_ratios)
